---
title: "credit card segmentation"
output:
  html_document: default
  word_document: default
  toc: TRUE
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Jeevan Hemmanna B
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())

x = c('ggplot2','corrgram','randomForest','unbalanced','C50','dummies','e1071','Information',
      'rpart','gbm','ROSE','DMwR','caret')

setwd("C:/Users/jhemmann/Desktop/Edwisor-project/")

lapply(x,require,character.only = TRUE)

rm(x)

library(cluster)
library(factoextra)
library(dplR)
library(reshape2)
library(tidyverse) ## manipulating and visualizing data (plyr, purrr, ggplot2, knitr...)
library(kableExtra) ## make nice tables with wrapper for kable()
library(cluster)    ## clustering algorithms and gap statistic
library(GGally) ## create matrix of variable plots
library(NbClust) ## clustering algorithms and identification of best K
library(varhandle) ## to create dummies 
library(ggplot2)
library(animation)
library(tidyr)
library(RColorBrewer)
library(dplyr)
```

## Problem Statement
This case requires trainees to develop a customer segmentation to define marketing strategy. The sample dataset summarizes the usage behaviour of about 9000 active credit card holders during the last 6 months. The file is at a customer level with 18 behavioural variables.

## Data PreProcessing
Here we are now doing a data preprocessing and we will be cresting the New KPI’s and also we will be doing data scaling and treating the outliers.

```{r,data_preProcessing}
set.seed(25)
credit =  read.csv('credit-card-data.csv',header = T , na.strings = c(" ","","NA","NAN"))

str(credit)

summary(credit)

missing_val = data.frame(apply(credit, 2, function(x){sum(is.na(x))}))
missing_val

credit$CREDIT_LIMIT[is.na(credit$CREDIT_LIMIT)] = median(credit$CREDIT_LIMIT, na.rm = T)
credit$MINIMUM_PAYMENTS[is.na(credit$MINIMUM_PAYMENTS)] = median(credit$MINIMUM_PAYMENTS, na.rm = T)
sum(is.na(credit))
```

## Deriving New KPI's
Here we are deriving New KPI as asked in a problem statements.

* Monthly average perchase is purchases by tenure
* monthly cash advance is cash advance by tenure
* creating categories by using One_off purchase and installment purchases
* limit usage is basically ratio of balance to limit usage.Lower limit usage implies cutomers are maintaing thier balance properly and it also mean that they have a  good credit score
* Pay_minPay is the ratio of payments to minimum payments

These are the new KPI's formed

```{r,derving_KPI}

#New Variables creation# 
credit$Monthly_Avg_PURCHASES <- credit$PURCHASES/(credit$TENURE)
credit$Monthly_CASH_ADVANCE <- credit$CASH_ADVANCE/(credit$TENURE)
credit$LIMIT_USAGE <- credit$BALANCE/credit$CREDIT_LIMIT
credit$MIN_PAYMENTS_RATIO <- credit$PAYMENTS/credit$MINIMUM_PAYMENTS

credit$PURCHASE_TYPE <- dplyr::case_when(
  credit$ONEOFF_PURCHASES == 0 & credit$INSTALLMENTS_PURCHASES == 0 ~ 'none',
  credit$ONEOFF_PURCHASES > 0 & credit$INSTALLMENTS_PURCHASES == 0 ~ 'oneoff',
  credit$ONEOFF_PURCHASES > 0 & credit$INSTALLMENTS_PURCHASES > 0 ~ 'both_oneoff_installment',
  credit$ONEOFF_PURCHASES ==0 & credit$INSTALLMENTS_PURCHASES > 0 ~ 'installment'
)
credit$PURCHASE_TYPE  <- as.factor(credit$PURCHASE_TYPE)

colnames(credit)

str(credit)

credit_copy = credit
```
## preparing for Machine Learning

Here we will be preparing the data for machine learning algorithm. we Need to convert the purchase_type column into binary variable.

```{r,Preparing_for_ML}

transformed_variables <- c("BALANCE","BALANCE_FREQUENCY", 
                           "PURCHASES", "ONEOFF_PURCHASES", "INSTALLMENTS_PURCHASES", 
                           "CASH_ADVANCE","PURCHASES_FREQUENCY","ONEOFF_PURCHASES_FREQUENCY",
                           "PURCHASES_INSTALLMENTS_FREQUENCY","CASH_ADVANCE_FREQUENCY", 
                           "CASH_ADVANCE_TRX", "PURCHASES_TRX", "CREDIT_LIMIT", 
                           "PAYMENTS", "MINIMUM_PAYMENTS","PRC_FULL_PAYMENT","TENURE",                                        "Monthly_Avg_PURCHASES","Monthly_CASH_ADVANCE",
                           "LIMIT_USAGE","MIN_PAYMENTS_RATIO") # vector of variables to be log transformed

credit_transformed_data <- credit %>%                           # preserve original dataset
  .[,-c(1,23)] %>%                                              #dropping CUST_ID and PURCHASE_TYPE
  mutate_at(vars(transformed_variables), funs(log(1 + .))) %>%  # add 1 to each value to avoid log0
  mutate_at(c(2:17), funs(c(scale(.))))                         # scale all numeric variables to mean of 0 & sd = 1
            
#taking only the variable which we dont use for KPI
credit_transformed_selected <- credit_transformed_data[,-c(1,3,6,17,14,15,16,13)]

colnames(credit_transformed_selected)

binary_purchase_type <- to.dummy(credit$PURCHASE_TYPE, "purchase_type")

credit_transformed_selected <- cbind(credit_transformed_selected,binary_purchase_type)

colnames(credit_transformed_selected)
```
```{r,corr_map}
cormat <- round(cor(credit_transformed_selected),3)
credit_corr <- melt(cormat)
```
```{r,fig.align='center'}
ggplot(data = credit_corr,aes (x = Var1,y = Var2,fill = value)) +
theme(axis.text.x = element_text(angle = 45, vjust = 1,size = 9, hjust = 1)) +
geom_tile()
```
By seeing the heatmap we can conclude that number of variable is too much and we need to reduce the number of variable.

## Standardrizing and scaling of the data

We can reduce the number of variable by using the PCA (principal Component Algorithm). For PCA to work we need to scale the data and then we will be using the PCA.
Before Applying the PCA we need to find the optimal value for number of component. we will be using the elbow method to find the optimal number of component

```{r,Data_scale}

credit_scale = scale(credit_transformed_selected)

credit_scale = as.data.frame(credit_scale)
dim(credit_scale)
```

```{r,PCA}
credit_pca <- prcomp(credit_scale)

names(credit_pca)

credit_pca$center
credit_pca$scale
credit_pca$rotation
dim(credit_pca$x)

std_dev <- credit_pca$sdev
pca_var <- std_dev^2

pca_varex <- pca_var/sum(pca_var)
```
```{r,fig.align='center'}
plot(pca_varex, xlab = "Principal Component",
   ylab = "Proportion of Variance Explained",
   type = "b")
```

This plot tells us that selecting 5 components we can preserve something around 87% of the total variance of the data. It makes sense, we’ll not use 100% of our variance, because it denotes all components, and we want only the principal ones. With this information in our hands, we can implement the PCA for 5 best components.

```{r,fig.align='center'}
plot(cumsum(pca_varex), xlab = "Principal Component",
   ylab = "Cumulative Proportion of Variance Explained",
   type = "b")
```

This plot shows that 5 components results in variance close to ~ 87%. Therefore, in this case, we'll select number of components as 5 [PC1 to PC5] and proceed to the modeling stage. This completes the steps to implement PCA on train data.
For modeling, we'll use these 5 components as predictor variables and follow the normal procedures.

```{r}
credit_pca_data <- data.frame(credit_pca$x)
credit_pca_data <- credit_pca_data[,1:5]
dim(credit_pca_data)
```
## Clustering Algorithm - K Means Clustering

Now Our data is ready after we have done the feature selection and scaling of the data.

Before we start the KMeans we need find the optimal cluster value.for the we will be elbow method to find the optimal number of cluster.
```{r,KMean}
kmean_withinss <- function(k) {
cluster <- kmeans(credit_pca_data, k)
return (cluster$tot.withinss)
}

max_k <- 10
wss <- sapply(2:max_k, kmean_withinss)

# Create a data frame to plot the graph
elbow <-data.frame(2:max_k, wss)
```
```{r,fig.align='center'}
ggplot(elbow, aes(x = X2.max_k, y = wss)) +
geom_point() +
geom_line() +
scale_x_continuous(breaks = seq(1, 20, by = 1))
```

From the graph, you can see the optimal k is 4, where the curve is starting to have a diminishing return. Once you have our optimal k, you run the algorithm with k equals to 4 and evaluate the clusters.

## Applying the K value 4

```{r,k_4}
pc_cluster_4 <-kmeans(credit_pca_data, 4)
cluster_df <- (pc_cluster_4$cluster)
pc_cluster_4$centers
pc_cluster_4$size

center <-pc_cluster_4$centers
center
# create dataset with the cluster number
cluster <- c(1: 4)
center_df <- data.frame(cluster, center)
# Reshape the data
center_reshape <- gather(center_df, features, values, PC1:PC5)
head(center_reshape)

cluster_scatter <- cbind(credit_pca_data,pc_cluster_4$cluster)
names(cluster_scatter)[6] <- "Cluster"
```
```{r,fig.align='center'}
ggplot(data=cluster_scatter, aes(x=PC1, y=PC2, color=factor(Cluster))) + 
geom_point()
```
From the scatter plot we can see that we have 4 cluster here and we each data point classified into different cluster.

To get a clear idea of how observation are classified into different cluster we will using the original data without the varibale which are used to create new KPI's and then adding the cluster classification which we get from the kMeans algorithm and then we will be grouping the data by cluster and we will be summarizing the other variable by mean and checking how well all the observation are classified.
```{r,insights_k_4}

col_kpi=c('CASH_ADVANCE_TRX', 'PURCHASES_TRX', 'Monthly_Avg_PURCHASES',
        'Monthly_CASH_ADVANCE', 'LIMIT_USAGE', 'MIN_PAYMENTS_RATIO')
cluster_4 <- credit[,col_kpi]
cluster_4 <- cbind(cluster_4,binary_purchase_type)
cluster_4 <- cbind(cluster_4,cluster_df)
cluster_4[1:10,]
dim(cluster_4)

cluster_group_by <- group_by(cluster_4,cluster_df)
cluster_summ <- summarise(cluster_group_by,CASH_ADVANCE_TRX = mean(CASH_ADVANCE_TRX),
                          PURCHASES_TRX= mean(PURCHASES_TRX),
                          Monthly_Avg_PURCHASES = mean(Monthly_Avg_PURCHASES),
                          LIMIT_USAGE = mean(LIMIT_USAGE),
                          Monthly_CASH_ADVANCE = mean(Monthly_CASH_ADVANCE),
                          MIN_PAYMENTS_RATIO = mean(MIN_PAYMENTS_RATIO),
                          purchase_type.both_oneoff_installment = mean(purchase_type.both_oneoff_installment),
                        purchase_type.installment = mean(purchase_type.installment),
                        purchase_type.none = mean(purchase_type.none),
                        purchase_type.oneoff = mean(purchase_type.oneoff))
cluster_summ <- t(cluster_summ)
colnames(cluster_summ) <- c("cluster-1","cluster-2","cluster-3","cluster-4")
cluster_summ
cluster_summ <- cluster_summ[-c(1,2,3),]

cluster_summ <- t(cluster_summ)
cluster_summ <- as.data.frame(cluster_summ)
cluster_summ$Monthly_Avg_PURCHASES <- log(cluster_summ$Monthly_Avg_PURCHASES)
cluster_summ$Monthly_CASH_ADVANCE <- log(cluster_summ$Monthly_CASH_ADVANCE)
cluster_summ <- t(cluster_summ)
cluster_summ <- as.data.frame(cluster_summ)
```
```{r,fig.align='center'}
barplot(as.matrix(cluster_summ), main="Cluster", ylab="values", beside=TRUE, 
      col=terrain.colors(5))
```
```{r}
cluster_count <- cluster_scatter %>%
                group_by(Cluster) %>%
                summarise(count_value = n())
cluster_count

cluster_percentage <- cluster_count %>%
                      mutate(percentage = (count_value/sum(count_value)*100))
cluster_percentage
```

**Some insight on the KMeans cluster-4**

from the above graph we can see that each cluster is clearly showing a distinguishing behaviour within the customers.

* Cluster 4 is the group of customers who have highest Monthly_avg purchases and doing both installment as well as one_off purchases, have comparatively good credit score. This group is about 31% of the total customer base.

* cluster 3 is taking maximum advance_cash and is paying comparatively less minimum payment and poor credit_score & doing no purchase transaction. This group is about 23% of the total customer base.

* Cluster 1 customers are doing maximum One_Off transactions and least payment ratio and credit_score on lower side This group is about 21% of the total customer base.

* Cluster 2 customers have maximum credit score and are paying dues and are doing maximum installment purchases. This group is about 25% of the total customer base

## Finding behaviour with 5 cluster

Even though we found that optimal value of K is 4 we will just see how the data will behave with number of cluster as 5.

For this cluster also we will be doing all the methods that are done for cluster 4 and finding how does the classification of the data point are done.

```{r,K_5}
pc_cluster_5 <-kmeans(credit_pca_data, 5)
cluster_df_5 <- (pc_cluster_5$cluster)
pc_cluster_5$centers
pc_cluster_5$size

center_5 <-pc_cluster_5$centers
center_5

cluster_5 <- c(1: 5)
center_df_5 <- data.frame(cluster_5, center_5)
# Reshape the data
center_reshape_5 <- gather(center_df_5, features, values, PC1:PC5)
head(center_reshape_5)

cluster_scatter_5 <- cbind(credit_pca_data,pc_cluster_5$cluster)
names(cluster_scatter_5)[6] <- "Cluster"
ggplot(data=cluster_scatter_5, aes(x=PC1, y=PC2, color=factor(Cluster))) + 
  geom_point()

col_kpi=c('CASH_ADVANCE_TRX', 'PURCHASES_TRX', 'Monthly_Avg_PURCHASES',
          'Monthly_CASH_ADVANCE', 'LIMIT_USAGE', 'MIN_PAYMENTS_RATIO')
cluster_5 <- credit[,col_kpi]
cluster_5 <- cbind(cluster_5,binary_purchase_type)
cluster_5 <- cbind(cluster_5,cluster_df_5)
cluster_5[1:10,]
dim(cluster_5)

cluster_group_by_5 <- group_by(cluster_5,cluster_df_5)
cluster_summ <- summarise(cluster_group_by_5,CASH_ADVANCE_TRX = mean(CASH_ADVANCE_TRX),PURCHASES_TRX= mean(PURCHASES_TRX),
                          Monthly_Avg_PURCHASES = mean(Monthly_Avg_PURCHASES),LIMIT_USAGE = mean(LIMIT_USAGE),Monthly_CASH_ADVANCE = mean(Monthly_CASH_ADVANCE),
                          MIN_PAYMENTS_RATIO = mean(MIN_PAYMENTS_RATIO),purchase_type.both_oneoff_installment = mean(purchase_type.both_oneoff_installment),
                          purchase_type.installment = mean(purchase_type.installment),purchase_type.none = mean(purchase_type.none),
                          purchase_type.oneoff = mean(purchase_type.oneoff))
cluster_summ <- t(cluster_summ)
colnames(cluster_summ) <- c("cluster-1","cluster-2","cluster-3","cluster-4","cluster-5")
cluster_summ
cluster_summ <- cluster_summ[-c(1,2,3),]

cluster_summ <- t(cluster_summ)
cluster_summ <- as.data.frame(cluster_summ)
cluster_summ$Monthly_Avg_PURCHASES <- log(cluster_summ$Monthly_Avg_PURCHASES)
cluster_summ$Monthly_CASH_ADVANCE <- log(cluster_summ$Monthly_CASH_ADVANCE)
cluster_summ <- t(cluster_summ)
cluster_summ <- as.data.frame(cluster_summ)

barplot(as.matrix(cluster_summ), main="Cluster", ylab="values", beside=TRUE, 
        col=terrain.colors(5))

cluster_count <- cluster_scatter_5 %>%
                group_by(Cluster) %>%
                summarise(count_value = n())
cluster_count

cluster_percentage <- cluster_count %>%
                      mutate(percentage = (count_value/sum(count_value)*100))
cluster_percentage
```
**Insights on cluster-5**

* we have a group of customers (cluster 5) having highest avergae purchases but there is Cluster 2 also having highest cash advance & secong highest purchase behaviour but their type of purchases are same.

* Cluster 4 and Cluster 2 are behaving similar in terms of Credit_limit and have cash transactions is on higher side.

**Since we cannot draw a proper conclusion with K value 5. We take K value 4 as optimal K and we will moving ahead considering K as 5 and drawing our marketing strategies.**

## Marketing strategies that can be drawn from above analysis

We Came to know that we have 4 kind of customers

**1. Group 1**

* This group is has minimum paying ratio and using card for just oneoff transactions (may be for utility bills only). This group seems to be risky group.

**2. Group 2**

* This group is performing best among all as cutomers are maintaining good credit score and paying dues on time. -- Giving rewards point will make them perform more purchases.

**3. Group 3**

* They have poor credit score and taking only cash on advance. We can target them by providing less interest rate on purchase transaction.


**4. Group 4**

* They are potential target customers who are paying dues and doing purchases and maintaining comparatively good credit score ) -- we can increase credit limit or can lower down interest rate -- Can be given premium card /loyality cards to increase transactions.

**Group 4 is the potential Target customer**